---
title: "explore_one_variable"
author: "vikram"
date: "24 January 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Explore One Variable

## Step 1: Load Diamonds dataset
```{r}
library('ggplot2')
summary(diamonds)
```

## Exploring the dataset
```{r}
head(diamonds)
?diamonds
```
# How many ordered factors in dataset?
```{r}
levels(diamonds$clarity)
levels(diamonds$cut)
levels(diamonds$color)
```

## If we check the output of above commands it makes sense to say that these are ordered, by also reading the description obtained by ?diamonds command

# Create a histogram of the price of
# all the diamonds in the diamond data set.

# =======================================
```{r}
qplot(data = diamonds, x = price)
```

# Price histogram summary
 
```{r}
summary(diamonds$price)
```

## Trying to explore the summary, as to which can be a correct central dependency measure
```{r}
qplot(data = diamonds, x = price) +
  scale_x_continuous(limits = c(0, 2000))

print(length(diamonds$price[diamonds$price < 2401])) #less than median
print(length(diamonds$price[diamonds$price > 3933])) #less than mean
```

## It seems the median accounts for the best result here.
# =====================================================================

# Explore the largest peak in the price histogram created earlier.

## Try limiting the x-axis, altering the bin width,
## and setting different breaks on the x-axis.

## Save images by using ggsave() command.
## ggsave() will save the last plot created.
## For example...
##                  qplot(x = price, data = diamonds)
##                  ggsave('priceHistogram.png')

## ggsave currently recognises the extensions eps/ps, tex (pictex),
## pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Cheaper diamonds
# ======================================================================
```{r}
qplot(data = diamonds, x = price, binwidth = 10) + 
  scale_x_continuous(limits = c(500, 1000))

qplot(data = diamonds, x = price, binwidth = 10,
      fill = I("#5670AB"), color = I('black')) +
  scale_x_continuous(limits = c(500, 1000))

qplot(data = diamonds, x = price, binwidth = 1,
      fill = I("#5670AB"), color = I('black')) +
  scale_x_continuous(limits = c(550, 650))

qplot(data = diamonds, x = price, binwidth = 1,
      fill = I("#5670AB"), color = I('black')) +
  scale_x_continuous(limits = c(600, 625))
```

# Price by cut histograms
## Break out the histogram of diamond prices by cut.

## You should have five histograms in separate
## panels on your resulting plot.

# ======================================================
## here, we find the summary of price/cut, easy to get the mean, median, etc
```{r}
by(diamonds$price, diamonds$cut, summary)
```
## Here, we find the what is highest price and which cut has it? so we progress slowly slowly by increasing the values until we are satisfied
```{r}
qplot(data = diamonds, 
      x = cut,
      y = price) +
  scale_y_continuous(limits = c(18750, 18830))
```
## Here, we find the what is lowest price and which cut has it?
```{r}
qplot(data = diamonds, 
      x = cut,
      y = price) +
  scale_y_continuous(limits = c(325, 330))
```

# 
```{r}
qplot(x = price, data = diamonds) + facet_wrap(~ cut, scales = 'free_y')
```

# 
```{r}
qplot(x = price/carat, data = diamonds, binwidth = 0.25, geom = 'histogram') + facet_wrap(~cut, scales = 'free_y') + 
  scale_x_log10()
```

## Investigate the price of diamonds using box plots,
## numerical summaries, and one of the following categorical
## variables: cut, clarity, or color.

# BOXPLOTS BY CUT
## =================================================================

```{r}
summary(diamonds)
levels(diamonds$cut)
qplot(data = diamonds, x = price, y = cut, geom = 'boxplot') +
  facet_wrap(~ cut)
```

# BOXPLOTS BY CLARITY
## =================================================================
```{r}
qplot(data = diamonds, x = clarity, y = price, geom = 'boxplot', color = clarity)
levels(diamonds$clarity)
?diamonds
```

# BOXPLOTS BY COLOR
## =================================================================
```{r}
qplot(data = diamonds, x = color, y = price, geom = 'boxplot', color = color) + 
  facet_wrap(~ color, scales = 'free_y')
```

# BOXPLOTS BY CUT
## =================================================================
```{r}
qplot(data = diamonds, x = cut, y = price/carat, geom = 'boxplot', color = cut) + 
  scale_y_continuous(limits = c(10000, 20000))
```
