---
title: "explore_many_variables"
author: "vikram"
date: "28 January 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(gridExtra)
```

## Explore the dataset
```{r}
pf = read.csv('/home/vikram.iyer/vikram-code-school/ND/P4/datasets/pseudo_facebook.tsv', sep = '\t', header = TRUE)
head(pf)
names(pf)
dim(pf)

```


## Third qualitative Variable
```{r}
p1 <- ggplot(aes(x = gender, y = age), data = subset(pf, !is.na(gender))) +
  geom_boxplot() 
```

```{r}
p2 <- ggplot(aes(x = gender, y = age), data = subset(pf, !is.na(gender))) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 5)

grid.arrange(p1, p2, ncol = 2)
```

```{r}
## lets compare the two using a line plot

ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median) +
  scale_x_continuous(limits = c(13,70), breaks = seq(0,70,5))

names(pf)
```

```{r}
library(dplyr)
pf.fc_by_age_gender <- subset(pf, !is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  arrange(age)

## the above code fails to work because,
# 1. we grouped by age first
# 2. we grouped by gender
# 3. now if without applying the ungroup, we are trying to arrange w.r.t age, but it is already grouped by gender, 
# so we need to remove one layer of grouping from gender and then apply arrange by age

## Now the below code will work
pf.fc_by_age_gender <- subset(pf, !is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)

## Reference: http://stackoverflow.com/questions/27207963/arrange-not-working-on-grouped-data-frame

##  we have used the subset() command, filter() can also be used
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)

pf.fc_by_age_gender
```

# Create a line graph showing the median friend count over the ages for each gender. Be sure to use the data frame you just created, pf.fc_by_age_gender
```{r}
names(pf.fc_by_age_gender)
ggplot(aes(x = age, y = median_friend_count), data = pf.fc_by_age_gender) +
  geom_line(stat = 'summary', aes(color = gender), fun.y = median)
```

```{r}
library(tidyr) ## reshape2 and tidyr both perform the operations that we desire, lets see examples of both
spread(subset(pf.fc_by_age_gender, select = c('gender', 'age', 'median_friend_count')), gender, median_friend_count)
suppressMessages(library(reshape2))
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, # dcast - dataframe, acast - array/matrix
                                  age ~ gender,
                                  value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)
```

# Plot the ratio of the female to male median friend counts using the data frame pf.fc_by_age_gender.wide
```{r}
ggplot(aes(x = age, y = female/male), data = pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(alpha = .3, yintercept = 1, linetype = 2)
```

# Third quantitative variable
## Create a variable called year_joined in the 'pf' data frame using the variable tenure and 2014 as the reference year
```{r}
#year_joined 
pf$year_joined <- floor(2014 - pf$tenure/365)
```

```{r}
summary(pf$year_joined)
table(pf$year_joined)
```
## use cut to separate out the year_joined var into 4 bins
## 2004-2009
## 2009-2011
## 2011-2012
## 2012-2014
```{r}
pf$year_joined.bucket <- cut(pf$year_joined,
                             c(2004,2009,2011,2012,2014))
pf
# Reference: https://www.r-bloggers.com/r-function-of-the-day-cut-2/
```

# Plotting 
```{r}
table(pf$year_joined.bucket, useNA = 'ifany')
ggplot(aes(x = age, y = friend_count), 
              data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

```{r}
ggplot(aes(x = age, y = friend_count), 
              data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', linetype = 4, fun.y = mean) # grand mean
```

# Friending rate solution
```{r}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```

# Create a line graph of mean of friendships_initiated per day (of tenure) vs. tenure colored by year_joined.bucket.
```{r}
ggplot(aes(x = tenure, y = friendships_initiated / tenure), 
              data = subset(pf, tenure >= 1)) + 
  geom_line(aes(color = year_joined.bucket), fun.y = mean)
```
## Here, we can say that, people who have more tenure, do not initiate many requests
```{r}
# we will apply some smoothers to reduce the noise
ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket), method = lm)

# Reference: http://scott.fortmann-roe.com/docs/BiasVariance.html
```

# explore yogurt.csv
```{r}
yo <- read.csv('/home/vikram.iyer/vikram-code-school/ND/P4/datasets/yogurt.csv')
names(yo)
summary(yo)
dim(yo)
head(yo)
?yo # no documentation
str(yo)
```

```{r}
## for now, convert id from int to factor
yo$id <- factor(yo$id)
str(yo)
```

```{r}
## plot histogram of yogurt prices
ggplot(aes(x = price), data = yo) +
  geom_histogram()
```

```{r}
#let's find the range of values of prices to plot it better
range(yo$price)
```

```{r}
## THis says that,we specify the limits to be from 20 to 70 for the prices
ggplot(aes(x = price), data = yo) +
  geom_histogram(fill = I('#F79420')) +
  scale_x_continuous(limits = c(20, 70), breaks = seq(20,70,2))
```

### As expected, this distribution has lots of discontinuity since the prices will not be in a continuous range
### Also, prices are mostly close to 20,30,40,50,60,70 


# No of purchases
## Create a new variable called all.purchases, which gives the total counts of yogurt for each observation or household.
## One way to do this is using the transform function. You can look up the function transform and run the examples of code at the bottom of the documentation to figure out what it does.

```{r}
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
## Insteadd of transform, even below code would have been fine, but a bit more verbose
### yo$all.purchases = yo$strawberry + yo$blueberry + yo$pina.colada + yo$plain + yo$mixed.berry
```

```{r}
## Lets create a histogram for this
ggplot(aes(x = all.purchases), data = yo, bindiwidth = 1) +
  geom_histogram(fill = I('#6C8129'))
```

# Create a scatterplot of price vs time. This will be an example of a time series plot. Resolve overplotting issues
```{r}
names(yo)
ggplot(aes(x = time, y = price), data = yo) +
    geom_jitter(alpha = 1/4, shape = 21, fill = I('833199'))
```

## We can say by looking at the above graph that, the prices have been increasing over time

# Sampling data
## When we have multiple observations to be taken care of, we must pick a random sample of smaller size and try exploring it and performing statistical observations
```{r}
set.seed(0009) # Set the seed for reproducible results
sample(levels(yo$id),16)
sample.ids <- sample(levels(yo$id),16)

ggplot(aes(x = time, y = price), data = subset(yo, id %in% sample.ids)) +
  facet_wrap( ~ id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 4)
```

# Scatterplot matrices

```{r}
pf[ , c(2:7)] 
```

# References
#### http://stackoverflow.com/questions/27207963/arrange-not-working-on-grouped-data-frame
#### https://s3.amazonaws.com/udacity-hosted-downloads/ud651/DataWranglingWithR.pdf
#### http://seananderson.ca/2013/10/19/reshape.html
#### http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/
#### https://www.r-bloggers.com/melt/
#### https://www.r-bloggers.com/r-function-of-the-day-cut-2/
#### http://scott.fortmann-roe.com/docs/BiasVariance.html
#### http://www.perossi.org/home/bsm-1
#### http://www.nuff.ox.ac.uk/Economics/papers/2013/Panel%20Data%20Demand%20-%20June%2020,%202013.pdf
#### http://htmlcolorcodes.com/ -> for hex color codes :D
#### http://www.statmethods.net/advgraphs/parameters.html
#### https://s3.amazonaws.com/udacity-hosted-downloads/ud651/scatterplotMatrix.pdf
#### http://www.statmethods.net/advgraphs/parameters.html