---
title: "explore_many_variables"
author: "vikram"
date: "28 January 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(gridExtra)
```

## Explore the dataset
```{r}
pf = read.csv('/home/vikram.iyer/vikram-code-school/ND/P4/datasets/pseudo_facebook.tsv', sep = '\t', header = TRUE)
head(pf)
names(pf)
dim(pf)

```


## Third qualitative Variable
```{r}
p1 <- ggplot(aes(x = gender, y = age), data = subset(pf, !is.na(gender))) +
  geom_boxplot() 
```

```{r}
p2 <- ggplot(aes(x = gender, y = age), data = subset(pf, !is.na(gender))) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 5)

grid.arrange(p1, p2, ncol = 2)
```

```{r}
## lets compare the two using a line plot

ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

```{r}
library(dplyr)
pf.fc_by_age_gender <- subset(pf, !is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  arrange(age)

## the above code fails to work because,
# 1. we grouped by age first
# 2. we grouped by gender
# 3. now if without applying the ungroup, we are trying to arrange w.r.t age, but it is already grouped by gender, 
# so we need to remove one layer of grouping from gender and then apply arrange by age

## Now the below code will work
pf.fc_by_age_gender <- subset(pf, !is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)

## Reference: http://stackoverflow.com/questions/27207963/arrange-not-working-on-grouped-data-frame

##  we have used the subset() command, filter() can also be used
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)

pf.fc_by_age_gender
```

# Create a line graph showing the median friend count over the ages for each gender. Be sure to use the data frame you just created, pf.fc_by_age_gender
```{r}
names(pf.fc_by_age_gender)
ggplot(aes(x = age, y = median_friend_count), data = pf.fc_by_age_gender) +
  geom_line(stat = 'summary', aes(color = gender), fun.y = median)
```

```{r}
library(tidyr) ## reshape2 and tidyr both perform the operations that we desire, lets see examples of both
spread(subset(pf.fc_by_age_gender, select = c('gender', 'age', 'median_friend_count')), gender, median_friend_count)
suppressMessages(library(reshape2))
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, # dcast - dataframe, acast - array/matrix
                                  age ~ gender,
                                  value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)
```

# Plot the ratio of the female to male median friend counts using the data frame pf.fc_by_age_gender.wide
```{r}
ggplot(aes(x = age, y = female/male), data = pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(alpha = .3, yintercept = 1, linetype = 2)
```

# Third quantitative variable
## Create a variable called year_joined in the 'pf' data frame using the variable tenure and 2014 as the reference year
```{r}
#year_joined 
pf$year_joined <- floor(2014 - pf$tenure/365)
```

```{r}
summary(pf$year_joined)
table(pf$year_joined)
```
## use cut to separate out the year_joined var into 4 bins
## 2004-2009
## 2009-2011
## 2011-2012
## 2012-2014
```{r}
pf$year_joined.bucket <- cut(pf$year_joined,
                             c(2004,2009,2011,2012,2014))
pf
# Reference: https://www.r-bloggers.com/r-function-of-the-day-cut-2/
```

# Plotting 
```{r}
table(pf$year_joined.bucket, useNA = 'ifany')
ggplot(aes(x = age, y = friend_count), 
              data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

```{r}
ggplot(aes(x = age, y = friend_count), 
              data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', linetype = 4, fun.y = mean) # grand mean
```

# Friending rate solution
```{r}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```

# Create a line graph of mean of friendships_initiated per day (of tenure) vs. tenure colored by year_joined.bucket.
```{r}
ggplot(aes(x = tenure, y = friendships_initiated / tenure), 
              data = subset(pf, tenure >= 1)) + 
  geom_line(aes(color = year_joined.bucket), fun.y = mean)
```
## Here, we can say that, people who have more tenure, do not initiate many requests
```{r}
# we will apply some smoothers to reduce the noise
ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket), method = lm)

# Reference: http://scott.fortmann-roe.com/docs/BiasVariance.html
```

# explore yogurt.csv
```{r}

```


# References
#### http://stackoverflow.com/questions/27207963/arrange-not-working-on-grouped-data-frame
#### https://s3.amazonaws.com/udacity-hosted-downloads/ud651/DataWranglingWithR.pdf
#### http://seananderson.ca/2013/10/19/reshape.html
#### http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/
#### https://www.r-bloggers.com/melt/
#### https://www.r-bloggers.com/r-function-of-the-day-cut-2/
#### http://scott.fortmann-roe.com/docs/BiasVariance.html