---
title: "facebook_data"
author: "vikram"
date: "21 January 2017"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(gridExtra)
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
#analyzing facebook data v1
```{r}
facebook_data = read.csv("~/vikram-code-school/ND/P4/datasets/pseudo_facebook.tsv", sep = '\t')
summary(facebook_data)
qplot(data = facebook_data, x = dob_day) +
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~ dob_month)
```

#analyzing facebook data v2
```{r}
pf = read.csv("~/vikram-code-school/ND/P4/datasets/pseudo_facebook.tsv", sep = '\t')
```
#plot friend_count
```{r}
qplot(data = pf, x = friend_count)
```
## the above code, will consider the outliers as it spans the entire data, but we can
## see that the major chunk of data is till 1000, so we may want to scale up things

# anyone of below works file
# xlim accepts a factor, with a starting and ending position
```{r}
qplot(data = pf, x = friend_count, xlim = c(0, 1000))
qplot(data = pf, x = friend_count) +
  scale_x_continuous(limits = c(0,1000))
```

# working with binwidth
```{r}
qplot(data = pf, x = friend_count, binwidth = 25) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0, 1000, 50))
```

# split the plot into male and female to see who has the max number of friends
```{r}
qplot(data = pf, x = friend_count) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_grid(. ~ gender)
```
## We see a new var called NA, which means people who were not registered with with either male of female, lets remove that
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender))) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_grid(. ~ gender)
```
## Done :)
## There is another function we could have used to omit na.
## data = na.omit(pf) could have also done the trick, however, all the NA's from the entire of ## dataset would be removed in that case, so if some one had 0 friends and that was not
## mapped, then that would be removed as well

# Count of female and male
```{r}
table(pf$gender)
```

# Stats by gender
```{r}
by(pf$friend_count, pf$gender, summary)
```
# Working with Tenure
```{r}
qplot(x = tenure/365, data = pf, binwidth = .25,
      color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7))
```

# The naming for xlab and ylab is not as good as it should be, lets change it
```{r}
qplot(x = tenure/365, data = pf, binwidth = .25,
      xlab = 'Years on FB',
      ylab = 'Sample number',
      color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7))
```

# Ages
```{r}
summary(pf)
qplot(data = pf, x= age, binwidth = 1,
      fill = I("#5670AB"), color = I('black')) + 
  scale_x_continuous(breaks = seq(0, 113, 5))
```

# transforming data, managing multiple histograms in one single plot
```{r}
p1 <- qplot(x = friend_count, data = pf)
p2 <- qplot(x = log10(friend_count + 1), data = pf)
p3 <- qplot(x = sqrt(friend_count), data = pf)
grid.arrange(p1,p2,p3,ncol=1)
```

# Do the above thing using scales
```{r}
p1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()

grid.arrange(p1, p2, p3, ncol = 1)
```

# Frequency polygons to overlay separate histograms into one

```{r}
qplot(data = subset(pf, !is.na(gender)), x = friend_count, binwidth = 10, 
      geom = 'freqpoly', color = gender) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50))
```

## the above plot makes it easier to compare the two (males and females). However, our question is to find who has the max number of friends. We might need to tweek this a bit more. Maybe change the y axis to something more meaningful

```{r}
qplot(data = subset(pf, !is.na(gender)), x = friend_count, binwidth = 10,
      y = ..count../sum(..count..),
      geom = 'freqpoly', color = gender, 
      xlab = 'Friend count',
      ylab = 'Proportion of users with that friend count') +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50))
```

## THis makes it much easier, let's try to find where the females beat males by changing the limits. :)
```{r}
qplot(data = subset(pf, !is.na(gender)), x = friend_count, binwidth = 10,
      y = ..count../sum(..count..),
      geom = 'freqpoly', color = gender, 
      xlab = 'Friend count',
      ylab = 'Proportion of users with that friend count') +
  scale_x_continuous(lim = c(300, 1000), breaks = seq(0, 1000, 50))
```


# References
# Facets: http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/
# Scales: http://docs.ggplot2.org/current/scale_continuous.html
# Themes: https://github.com/jrnold/ggthemes#examples
# More Themes: http://docs.ggplot2.org/0.9.2.1/theme.html 
# GridExtra: http://lightonphiri.org/blog/ggplot2-multiple-plots-in-one-graph-using-gridextra
# Linear Regression: https://en.wikipedia.org/wiki/Linear_regression#Assumptions
# Normal Distribution: https://en.wikipedia.org/wiki/Normal_distribution
# Log Transformations: https://www.r-statistics.com/2013/05/log-transformations-for-skewed-and-wide-distributions-from-practical-data-science-with-r/