---
title: "facebook_data"
author: "vikram"
date: "21 January 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
#analyzing facebook data v1
```{r}
facebook_data = read.csv("~/vikram-code-school/ND/P4/datasets/pseudo_facebook.tsv", sep = '\t')
summary(facebook_data)
qplot(data = facebook_data, x = dob_day) +
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~ dob_month)
```

#analyzing facebook data v2
```{r}
pf = read.csv("~/vikram-code-school/ND/P4/datasets/pseudo_facebook.tsv", sep = '\t')
```
#plot friend_count
```{r}
qplot(data = pf, x = friend_count)
```
## the above code, will consider the outliers as it spans the entire data, but we can
## see that the major chunk of data is till 1000, so we may want to scale up things

# anyone of below works file
# xlim accepts a factor, with a starting and ending position
```{r}
qplot(data = pf, x = friend_count, xlim = c(0, 1000))
qplot(data = pf, x = friend_count) +
  scale_x_continuous(limits = c(0,1000))
```

# working with binwidth
```{r}
qplot(data = pf, x = friend_count, binwidth = 25) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0, 1000, 50))
```

# split the plot into male and female to see who has the max number of friends
```{r}
qplot(data = pf, x = friend_count) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_grid(. ~ gender)
```
## We see a new var called NA, which means people who were not registered with with either male of female, lets remove that
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender))) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_grid(. ~ gender)
```
## Done :)
## There is another function we could have used to omit na.
## data = na.omit(pf) could have also done the trick, however, all the NA's from the entire of ## dataset would be removed in that case, so if some one had 0 friends and that was not
## mapped, then that would be removed as well

# Count of female and male
```{r}
table(pf$gender)
```

# Stats by gender
```{r}
by(pf$friend_count, pf$gender, summary)
```
# Working with Tenure
```{r}
qplot(x = tenure/365, data = pf, binwidth = .25,
      color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7))
```

# The naming for xlab and ylab is not as good as it should be, lets change it
```{r}
qplot(x = tenure/365, data = pf, binwidth = .25,
      xlab = 'Years on FB',
      ylab = 'Sample number',
      color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7))
```

# Ages
```{r}
summary(pf)
qplot(data = pf, x= age, binwidth = 1,
      fill = I("#5670AB"), color = I('black')) + 
  scale_x_continuous(breaks = seq(0, 113, 5))
```

# transforming data, managing multiple histograms in one single plot
```{r}
p1 <- qplot(x = friend_count, data = pf)
p2 <- qplot(x = log10(friend_count + 1), data = pf)
p3 <- qplot(x = sqrt(friend_count), data = pf)
grid.arrange(p1,p2,p3,ncol=1)
```

# Do the above thing using scales
```{r}
p1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()

grid.arrange(p1, p2, p3, ncol = 1)
```

# References
# Facets: http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/
# Scales: http://docs.ggplot2.org/current/scale_continuous.html
# Themes: https://github.com/jrnold/ggthemes#examples
# More Themes: http://docs.ggplot2.org/0.9.2.1/theme.html 