---
title: 'problem_set: explore two variables'
author: "vikram"
date: "28 January 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## explore the dataset
```{r}
library(ggplot2)
suppressPackageStartupMessages(library(ggplot2))
?diamonds
names(diamonds)
```

```{r}
dim(diamonds)
```

```{r}
levels(diamonds$cut)
```
```{r}
levels(diamonds$color)
```
```{r}
levels(diamonds$clarity)
```

# task 1
## create a scatterplot of price vs x. using the ggplot syntax.

```{r}
ggplot(aes(x = price, y = x), data = diamonds) +
  geom_point()
```

```{r}
# is there anything strange about the scatterplot?
# 1. outliers, some entries with length = 0mm and still a price associated with it
# 2. exponential growth of price with length
```

```{r}
# corelation between price and x, price and y, price and z
cor.test(diamonds$price, diamonds$x)
cor.test(diamonds$price, diamonds$y)
cor.test(diamonds$price, diamonds$z)
```

# task 2
## create a scatterplot of price vs depth
```{r}
## also add a transparency factor of 1/100 i.e. 0.01 and add breaks of 2units on y-axis
ggplot(aes(x = price, y = depth), data = diamonds) +
  geom_point(alpha = 0.01) +
  geom_smooth(method = lm, color = 'red', linetype = 2) +
  scale_y_continuous(breaks = seq(0,80,2))

## looking at the graph, what is range of depth between which most of the values lie?
## 60 to 64
```

```{r}
## corelation of depth vs price
cor(diamonds$depth, diamonds$price)
with(diamonds, cor.test(price, depth))

```

```{r}
## can we predict price based on depth?
## No. Because, the value is very much close to 0 which is not a strong mark for prediction, we need something above 70% i.e. (<= -.7 or >= +.7)
```


# task 3
## Create a scatterplot of price vs carat and omit the top 1% of price and carat values
```{r}
ggplot(aes(x = carat, y = price), data = diamonds) +
  xlim(0, quantile(diamonds$carat,0.99)) +
  ylim(0, quantile(diamonds$price,0.99)) +
  geom_point()
```

# tassk 4
## create a scatterplot of price vs. volume (x*y*z), This is a very rough approximation for a diamond's volume
```{r}
ggplot(aes(x = (x*y*z), y = price), data = diamonds ) +
  geom_point()
```
## Create a new variable for volume in the diamonds data frame (will be used later)
```{r}
diamonds$volume <- (diamonds$x*diamonds$y*diamonds$z)
```

```{r}
## observations from price vs volume scatterplot
## 1. outliers: volume = 0
sum(diamonds$volume == 0)
```

```{r}
## lets find whiich are they
diamonds[diamonds$volume == 0,]
```

```{r}
## corelation between price and volume
cor.test(diamonds$price, diamonds$volume)
## 0.9 seems to be a very high value for the corelation coefficient, even if we ignore the 20 outliers, it looks quite something which we can bet upon
```

```{r}
## lets form subset where the volume is not 0
new_diamonds <- subset(diamonds, volume != 0)

## now lets find th corelation
cor.test(new_diamonds$price, new_diamonds$volume)

## So it seems there is some corelation and we can predict the prices from the volume
```


