---
title: 'problem_set: explore two variables'
author: "vikram"
date: "28 January 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## explore the dataset
```{r}
library(ggplot2)
suppressPackageStartupMessages(library(ggplot2))
?diamonds
names(diamonds)
```

```{r}
dim(diamonds)
```

```{r}
levels(diamonds$cut)
```
```{r}
levels(diamonds$color)
```
```{r}
levels(diamonds$clarity)
```

# task 1
## create a scatterplot of price vs x. using the ggplot syntax.

```{r}
ggplot(aes(x = price, y = x), data = diamonds) +
  geom_point()
```

```{r}
# is there anything strange about the scatterplot?
# 1. outliers, some entries with length = 0mm and still a price associated with it
# 2. exponential growth of price with length
```

```{r}
# corelation between price and x, price and y, price and z
cor.test(diamonds$price, diamonds$x)
cor.test(diamonds$price, diamonds$y)
cor.test(diamonds$price, diamonds$z)
```

# task 2
## create a scatterplot of price vs depth
```{r}
## also add a transparency factor of 1/100 i.e. 0.01 and add breaks of 2units on y-axis
ggplot(aes(x = price, y = depth), data = diamonds) +
  geom_point(alpha = 0.01) +
  geom_smooth(method = lm, color = 'red', linetype = 2) +
  scale_y_continuous(breaks = seq(0,80,2))

## looking at the graph, what is range of depth between which most of the values lie?
## 60 to 64
```

```{r}
## corelation of depth vs price
cor(diamonds$depth, diamonds$price)
with(diamonds, cor.test(price, depth))

```

```{r}
## can we predict price based on depth?
## No. Because, the value is very much close to 0 which is not a strong mark for prediction, we need something above 70% i.e. (<= -.7 or >= +.7)
```


# task 3
## Create a scatterplot of price vs carat and omit the top 1% of price and carat values
```{r}
ggplot(aes(x = carat, y = price), data = diamonds) +
  xlim(0, quantile(diamonds$carat,0.99)) +
  ylim(0, quantile(diamonds$price,0.99)) +
  geom_point()
```

# tassk 4
## create a scatterplot of price vs. volume (x*y*z), This is a very rough approximation for a diamond's volume
```{r}
ggplot(aes(x = (x*y*z), y = price), data = diamonds ) +
  geom_point()
```
## Create a new variable for volume in the diamonds data frame (will be used later)
```{r}
diamonds$volume <- (diamonds$x*diamonds$y*diamonds$z)
```

```{r}
## observations from price vs volume scatterplot
## 1. outliers: volume = 0
sum(diamonds$volume == 0)
```

```{r}
## lets find whiich are they
diamonds[diamonds$volume == 0,]
```

```{r}
## corelation between price and volume
cor.test(diamonds$price, diamonds$volume)
## 0.9 seems to be a very high value for the corelation coefficient, even if we ignore the 20 outliers, it looks quite something which we can bet upon
```

```{r}
## lets form subset where the volume is not 0 and not greater than 800
new_diamonds <- subset(diamonds, volume != 0 & volume < 800)

## now lets find th corelation
cor.test(new_diamonds$price, new_diamonds$volume)

## So it seems there is some corelation and we can predict the prices from the volume
```

```{r}
## adjustments in price vs volume
ggplot(aes(x = volume, y = price), data = new_diamonds) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = lm)
```

# task 5
##  Use the function dplyr package to create a new data frame containing info on diamonds by clarity
```{r}
library(dplyr)
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n = n())

diamondsByClarity
```
# summary data frames with the mean price by clarity and color are created below.
```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
```


## write additional code to create two bar plots on one output image using the grid.arrange() function from the package gridExtra
```{r}
p1 <- ggplot(diamonds_mp_by_clarity, aes(x = clarity, y = mean_price, fill = clarity)) +
  geom_bar(stat = "summary")
p1
```

```{r}
## ctd.
p2 = ggplot(diamonds_mp_by_color, aes(x = color, y = mean_price, fill= color)) +
  geom_bar(stat = "summary")
p2
```

```{r}
## ctd
library(gridExtra)
grid.arrange(p1,p2, ncol =2)
```

```{r}
# mean price for s12 is highest while it is lowest for VVS1
# mean price for color 'J' is highest while it is lowest for E
# in short: Mean price tends to decrease as clarity improves. The same can be said for color.
```


# References: 
#### http://www.ats.ucla.edu/stat/r/faq/smooths.htm
#### http://simplystatistics.org/
#### https://en.wikipedia.org/wiki/Local_regression
#### http://www.mindmapinspiration.com/wp-content/uploads/2009/02/curious-brain-mind-map-10.jpg
#### http://flowingdata.com/2014/02/27/how-to-read-histograms-and-use-them-in-r/
#### http://docs.ggplot2.org/0.9.3/geom_bar.html
#### https://rstudio-pubs-static.s3.amazonaws.com/52545_d5a11147091d477eb66cb9d22ea649a2.html -> used only for visual improvements like filling color, etc

